### 2015-07-07 abstract.tex
initial commit of soro paper on automous grasping

This paper presents the development of an autonomous motion planning algorithm for a soft planar grasping manipulator capable of grasp-and-place operations by encapsulation.
The gripping end of the soft manipulator is fabricated without weakening seams using lost-wax casting instead of the commonly used multi-layer lamination process.
The soft manipulation system can grasp randomly positioned objects within its reachable envelope and move them to a desired location without human intervention.
The autonomous planning system leverages the compliance and continuum bending of the soft grasping manipulator to achieve repeatable grasps in the presence of uncertainty.
Modeling uncertainties arise from a simplifying constant-curvature assumption, unrepresented manipulator dynamics, stick-slip friction, and  non-linear fluidic control. 
These uncertainties are compensated by the inherent compliance of the soft gripper design and the motion planning strategy.
Additionally, when fully actuated, the manipulator remains compliant allowing it to conform to un-modeled object geometries.
A suite of experiments is presented that demonstrates the system's capabilities.