### 2015-07-07 introduction.tex

\section{Introduction}

%High Level Vision
\rkk{
Soft robots exhibit continuum body motion, large scale deformation, high compliance and adjustable impedance  compared to traditional rigid-bodied robots with high impedance \cite{trivedi2008soft}. 
% whereas humans and animals have extremities with variable impedance.
%Soft robots won't necessarily replace existing rigid robot technologies, but complement it, because they can enable static and dynamic manipulations systems in nature are capable of.
%A further motivating factor for this work is that soft robots are inherently safer for interaction with humans and the environment due to their structural compliance and they can be lighter, one of the reasons why we research on that path.
%Fluidic robots have uses in scenarios where a classical robot would cause sparks and are dangerous. Why do we need accuracy if we have compliance. A 3DOF gripper is inherently not soft. A rigid gripper can not manipulate delicate objects without a lot of sensing.
}
Such characteristics make this class of robots well-suited for highly dexterous tasks and interactions that require conformation to environmental uncertainty.

Our goal is to develop a soft planar fluid-powered gripper and a motion planning algorithm that leverages a soft morphology to robustly grasp, drag and place objects. 
%By utilizing softness in the design of our robot, we gain the ability to adapt to environmental variation \cite{mcmahan2006field}, harmlessly collide with the environment \cite{marchese2014whole}, and build resilience to unexpected external loads \cite{tolley2014resilient};

\begin{figure}[htb]
\centering
   \includegraphics[width=0.9\columnwidth]{Figures/experimental_results/egg_approach/egg_approach_sequence}
   \caption{\rkk{The soft manipulator is grasping a hollowed-out egg. The robot repeatably approached, grasped and moved this delicate object without breaking it.}}
   \label{fig:egg_approach_sequence}
\end{figure}

%Challenges 
Our manipulator is highly compliant and its motion is not as precise as more traditional rigid-bodied robots. 
The control of its configuration is not only limited by the compliance of its low-pressure actuation, but also by its inherent elasticity.
%Additionally, the manipulator's compliant nature complicates the process of determining and sensing the forward kinematics.
Despite the constraints on accurate positioning introduced by the softness of the arm, we can still robustly execute autonomous grasping tasks.

%\begin{figure}[htp]
%   \centering
%   \includegraphics[width=0.7\columnwidth]{Figures/introduction/ExperimentalSetup}
%   \caption{A soft robotic arm with gripper autonomously approaching and grasping an object on a plane.}
%   \label{fig:intro}
%\end{figure}

%Our Approach
%To address these challenges, 
%\rkk{We developed planning and control algorithms for a soft robotic arm with gripper.}
\rkk{
The fluid-powered gripper at the end of the arm can grasp an object through an open-loop controlled bending motion, even if the gripper is positioned relatively inaccurately in relation to the object to be grasped.
%The gripper can conform due to his compliance to variations in object geometry, allowing it to successfully grasp.
}
The design of the gripper itself is inspired by work of Polygerinos et al.~\cite{polygerinos2013towards}
The design is advantageous for grasping, because it exhibits high curvature, minimal radial expansion, and remains compliant during actuation. 
We can repeatably fabricate the gripper using a lost-wax casting process instead of the commonly used soft lithography technique, a multi-layer lamination process.
Without weakening seams, the gripper is not prone to de-lamination under high deformations. 
By abandoning the need for a lamination process, arbitrary shaped internal channels can be achieved. 
The gripper was used as a manually actuated single finger with a bend sensor as part of a 3-fingered-hand in \cite{homberg2015haptic}.

We attach the gripper to a multi-segment soft manipulator to enable autonomous grasp-and-place capabilities on a plane. 
Positional feedback is provided in real-time from a camera system.
%and actuation of the gripper and manipulator is driven by an external fluidic cylinder array.
We present a planning algorithm that advances the arm through all necessary states of the grasp-and-place operation.
A minimal strain, collision-free approach to the object of interest is found by posing the plan as a series of constrained nonlinear optimization problems.
The system first plans concentric approach circles shrinking from the initial end-effector pose down to the object diameter.
Next, the system searches for locally optimal manipulator configurations that constrain the end-effector to lie on these approach circles, so that the manipulator does not collide with the object. 
The manipulator is then moved between these plans using closed-loop trajectories. 
After successfully approaching the object, the gripper encapsulates it.
We experimentally validate the system's ability to repeatably and autonomously grasp-and-place a randomly placed object.

\subsection{Related Work}
There are several examples of soft grippers described in recent literature, \rkk{we will mainly focus on fluidic-based systems.} 
Deimel and Brock \cite{deimel2013compliant} developed a pneumatically actuated three-fingered hand made of reinforced silicone that is mounted to a hard robot and capable of grasping.
More recently, they have developed an anthropomorphic soft pneumatic hand capable of dexterous grasps, which is not mounted to a robot, but instead just held by a human\cite{deimel2014novel}. 
Ilievski et al. \cite{ilievski2011soft} create \rkk{through a soft-lithography fabrication} a pneumatic starfish-like gripper composed of \rkk{an array of silicone chambers and a PDMS membrane.} 
\rkk{
The gripper is hanging on a string and grasps objects like an egg or a mouse in an open-loop controlled manner.
} 
Stokes et al. \cite{Stokes2014hybrid} use a soft elastomer quadrupedal robot to grasp objects in a hard-soft hybrid robotic platform. 
A puncture resistant soft pneumatic gripper is developed by Shepherd et al. in \cite{shepherd2013soft}. 
An alternative to positive pressure actuated soft grippers is a robotic gripper that makes use of granular material jamming developed by Brown et al. and detailed in \cite{brown2010universal}.
\rkk{The soft octopus-inspired arms developed in \cite{calisti2010study} and \cite{calisti2011octopus} are not fluidic powered, but instead use cables to pull rigid fixtures embedded within an elastomer body. 
The arms were capable of grasping objects like pens or screws.}
A soft robotic tentacle developed in \cite{martinez2013robotic} was able to hold a flower and a horseshoe-shaped object.
The closest related soft pneumatic actuator design to our current work is the fast Pneu-net designs by Mosadegh et al. detailed in \cite{mosadegh2014pneumatic} and by Polygerinos et al. detailed in \cite{polygerinos2013towards}.
These finger-like actuators deform with minimal volume change and can bend to high curvatures.
\rkk{None of the above described grippers were controlled autonomously to perform their tasks and accordingly no statement on repeatability of the autonomous execution was given.}

\rkk{Cho et al. \cite{cho2009review} review several manufacturing processes for soft biomimetic robots. 
The vast majority of soft elastomer robots rely on the processes of soft lithography \cite{xia1998soft} and/or shape deposition manufacturing \cite{cham2002fast}. 
Especially noteworthy is the use of lost-wax fabrication for jammable skin chambers, which get vacuumed to stiffen up \cite{steltz2009soft}. 
The gripper presented here also uses a lost-wax molding technique, but with a different type of actuation in mind: the obtained cavity structures are inflated to cause the gripper to bend.}

%More generally, the literature has several examples of soft pneumatic elastomer robots. 
%Many of these robots use open-loop controllers. 
%There are soft fluid-powered rolling robots \cite{correll2010soft, onal2011soft, marchese2011soft}. 
%Also, there are soft-bodied robotic fish powered by pneumatics \cite{marchese2014autonomous} and hydraulics \cite{katzschmann2014hydraulic}. 
%There are soft legged pneumatic walkers \cite{shepherd2011multigait, tolley2014resilient}, explosive jumpers \cite{shepherd2013using}, and 
More recently, Marchese et al. demonstrated closed-loop position control of multi-segment soft planar fluidic elastomer manipulators in free space \cite{marchese2014design} and in confined spaces \cite{marchese2014whole}. Positioning is achieved using forward and inverse kinematic models and real-time camera measurements. 
The manipulator curvatures are controlled in real-time using a curvature-volume cascaded control law.
\rkk{The work presented in this paper builds on top of this system and expands its capabilities.
The previous version was only suitable for contactless motions in space, that would be suitable for mere inspection.
The work presented here is capable of autonomous object manipulation on a plane. 
Furthermore, this work shows that combining two different soft actuator types can leverage the strength of each and together achieve a more complicated task of manipulating objects.
The control of the arm is improved from the previous version by using a more robust vision tracking approach. 
Previously, there were only single markers along the arm, which worked for a limited motion range, but was prone to fail if the markers go out of sight or another object is introduced to the environment. 
Also, many trial runs were needed to arrive at the various final positions in the maze.
%, so the previous system was no repeatable and robust. 
}

\subsection{Contributions}
This work differs from the previous work in that we %create an entirely soft and autonomously controlled \emph{grasping} manipulation system. 
do not attach a novel soft gripper to a hard arm nor do we move the gripper around manually, but we rather take on the challenge of grasping-and-placing objects with a seven degrees of freedom planar arm made entirely from soft rubber. 
We provide in this work the following contribution to soft robotics:
\begin{itemize}
  \item \rkk{A soft manipulator consisting of a multi-segment arm with a gripper to enable fully compliant planar grasping.}
  \item \rkk{A planning algorithm to grasp-and-place randomly positioned objects on a planar surface using a 7 DOF soft manipulator.}  
  \item \rkk{Repeatable successful grasping demonstrations with a physical prototype and an experimental characterization of the uncertainty regions that can be tolerated by the soft gripper.}
  %\item \rkk{End-effector control of the grasping manipulator attached to a soft arm using a motion tracking system}.
  \item \rkk{Demonstrate that soft robots do neither require force sensing nor accurate positioning to allow for proper manipulation of delicate objects.} 
\end{itemize}

%This paper is organized as follows: In Section~\ref{sec:System_Overview} we provide an overview of the planar manipulation system. In Section~\ref{sec:soft_grasping_manipulator} the gripper morphology is described, its operating principles are discussed in relation to existing morphologies, and the lost wax fabrication process is explained. Section~\ref{sec:processing_and_control} details the planning algorithms for grasp-and-place. Section~\ref{sec:experimental_results} shows results from experimental runs and provides evaluations and experimental insights. Lastly, Section~\ref{sec:conclusion} concludes by providing a summary and discussion.