### 2015-10-08 experimental_results.tex
final adjustements before revision submission

\section{Experimental Results}
\label{sec:experimental_results}
In the following section summarizes the results from experiments with the soft manipulator. We discuss the grasping of delicate objects and cylindrical objects as well as the repeatability and success rate of the autonomous system.

\subsection{\rkk{Grasping Delicate Objects}}
The soft manipulator can pick up eggs and other delicate objects without breaking them. 
Figure~\ref{fig:egg_approach_sequence} shows how the manipulator approaches and grasps an egg.
Delicate objects can be manipulated without requiring a shape or a force sensor within its structure, since the compliant gripper body conforms to the object.
Rigid-bodied grippers usually rely on force sensing or another type of sensory feedback to avoid damage caused to the object.

\begin{figure}[htb]
\centering
   \includegraphics[width=0.8\columnwidth]{Figures/experimental_results/uncertainty}
   \caption{Experimental characterization of the gripper's capture region: the allowable positioning uncertainty is determined through repeated placements of the center of a cylindrical object at different points on a grid relative to the gripper. Blue dots indicate all object center positions for which a grasp could be performed successfully, red crosses show the positions where a grasp failed. The grey line outlines an area for the object to be positioned within so the gripper can grasp it. The evaluation of the capture region was performed similarly to a method described in \cite{dogar2010push}.}
   \label{fig:grasp_uncertainty}
\end{figure}

\subsection{\rkk{Repeated Grasp Experiments}}
We performed 25 experimental grasp-and-place trials at various positions to demonstrate the capabilities and repeatability of our system.
The results of all the trials are shown in Figure~\ref{fig:allTestsOverlaid}.
One representative approach, grasp, and retract move is shown in Figure~\ref{fig:oneTrialVsTime}.
In 23 of 25 experimental trials, the manipulator successfully achieved the task of grasping an object and bringing it to a bin \rkk{location shown in red}.
The object was placed five times on each of the five points marked on the board.
The markers only serve as a reference point for the user to place the object roughly at the same point at every repetition.
The user's placing accuracy is not important to the algorithm, the tracking system newly registers every single time the position of the placed object.
The five points were chosen to approximately represent the major portion of the manipulator's reachable workspace.
As long as the root of the gripper stops so that the object is located within the capture region, the gripper will pick it up through its sweeping closing motion. The capture region is shown in grey in Figure~\ref{fig:grasp_uncertainty}. 
The evaluation of the capture region is performed similarly to a method described in Dogar and Srinivasa\cite{dogar2010push} work on determining capture regions for a push-grasp of a classical robotic gripper.
Despite positioning inaccuracies of the soft manipulator, the gripper can nevertheless successfully perform a grasp of an object.

When the arm reaches the red bin location, it drops the object.

\begin{figure*}[htbp]
\begin{centering}
  \includegraphics[width=0.5\textwidth]{Figures/experimental_results/allTestsOverlaid.eps}
\includegraphics[width=0.385\textwidth]{figures/experimental_results/five_grasps.jpg}
\caption[Left: All 25 experimental grasping trials.]{Complete set of experimental grasp-and-place trials. In these experiments, the arm moves from an initial straightened configuration to grasp a round object placed in one of five locations (A-E). The arm then returns the object to a bin location shown in red. For each trial, a seven degrees of freedom manipulator representation is generated at both the \emph{grasped} and \emph{released} state using experimental data and is shown in blue. The corresponding 1 degree-of-freedom end-effector representation is shown in black. The round object's measured position at each state is shown in green. Right: Overlaid photographs of the manipulator grasping an object placed at each of the five locations. }\label{fig:allTestsOverlaid}
\end{centering}
\end{figure*}

\rkk{The unsuccessful trials happened due to stick-slip friction between the roller bearings and the table surface.
Our kinematic modeling does not account for this non-linear behavior, which acts as a disturbance and can lead to failure to arrive at the next waypoint.
To be more specific, in one of the trials, the robot slowed down too much before it almost reached its next waypoint and because of friction the arm got to a full stop. 
The proportional gain of the curvature controller was not able to compensate for that small positional delta and since the relatively low saturation level of the integrator portion of the controller was saturated, the arm did not move to the next waypoint pose. 
It is to note, that the saturation level for the integrator is defined by a safety limit on the maximal inflation of the robotic arm.
In the other unsuccesful trial, the stick-slip friction also caused the arm to halt before a waypoint. The controller built up enough forcing due to inflation, so that the arm slipped over the close waypoint without having all of its single arm segment curvature within an acceptable epsilon.
The controller then tried to swing back to fulfill the missed waypoint, missed it again and that finally caused the whole arm to oscillate back-and-forth and eventually push the object off the table.
In one of the other trials, the grasp and return was successfully performed, but a small overshoot over the final bin location at the end of the return caused the arm to drop off the table, which could have been avoided if the table would have not been too small towards the right.
This outlier is shown in Figure~\ref{fig:allTestsOverlaid}.
Overall, the experiments show that the system was repeatably able to autonomously locate a randomly placed object within its workspace, plan the arm motions, and perform the task of grasping and placing the object.}

\subsection{\rkk{Experimental Insights and Limitations}}
The system can drag payloads of less than 40\unit{g}, higher payloads cause the cylindrical arm segments to stall and possibly lift off the table without moving the payload.
There is a trade-off between the reachable workspace and the maximum payload.
As the length of the arm increases, more workspace can be reached while less payload can be manipulated. 
Most of the payload capability is already used up by the attached gripper itself.
A smaller gripper would allow for larger payloads to get picked up, but consequently only smaller objects can be grasped.
The workspace of the manipulator is limited to the top and left by the maximal extension length of the arm, and to the bottom by the maximum bending curvature, which the arm can achieve without over-actuating a single segment.
Objects were only grasped within the left quadrant of the arm, because of the gripper orientation and an upright initial starting pose.

\begin{figure*}[htbp]
\begin{centering}
  \includegraphics[width=0.375\textwidth]{Figures/experimental_results/oneTrialVsTime.eps}
\includegraphics[width=0.30\textwidth]{figures/experimental_results/exp5a-4_sweep_to_object.jpg}
\includegraphics[width=0.31\textwidth]{figures/experimental_results/exp5a-4_sweep_to_bin.jpg}
  \caption{Left: A time series representation of an experimental grasp-and-place trial for an object located at point E. Here, the locally optimal planned manipulator configurations as well as planned sequential approach circles are shown as black dotted curves. The arm and gripper are shown in their experimentally determined configuration representations at 1 second intervals. The cyan configurations represent the manipulator prior to grasping the object, that is moving from its initial configuration to the object's location. The magenta configurations represent the manipulator after grasping the object, that is moving from the object's location back to the bin location shown in red. Right: A photograph of the arm moving from its initial pose to the object and from the object to the release location, respectively. }
\label{fig:oneTrialVsTime}
\end{centering}
\end{figure*}