### 2015-10-15 abstract.tex
all changes to the intro based on daniela's input.

This paper presents the development of an autonomous motion planning algorithm for a soft planar grasping manipulator capable of grasp-and-place operations by encapsulation
\dr{under uncertainty in the position and shape of the object.}
The gripping end of the soft manipulator is fabricated without weakening seams using lost-wax casting instead of the commonly used multi-layer lamination process.
The soft manipulation system can grasp randomly positioned objects within its reachable envelope and move them to a desired location without human intervention.
The autonomous planning system leverages the compliance and continuum bending of the soft grasping manipulator to achieve repeatable grasps in the presence of uncertainty.
A suite of experiments is presented that demonstrates the system's capabilities.