### 2015-10-15 processing_control.tex
implements proposed changes by Daniela about the processing section

\section{Planning and Control}
\label{sec:processing_and_control}
This section covers our approach to preplanning motion waypoints for the soft robot and controlling the manipulator along those points. 
%We cover existing procedures reused within the system, and detail the system's new architecture for performing autonomous grasp-and-place tasks, including its main component, the Grasp Object Planner.

\subsection{Kinematic Control}
The forward kinematics algorithm \texttt{forwKin()} assumes piece-wise constant curvature according to \cite{webster2010design} and \cite{marchese2014design}. 
%This algorithm was experimentally validated for the soft planar arm in \cite{marchese2014design}.
In order to uniquely fit a configuration representation to measured endpoint data in real-time, we use a previously developed single segment inverse kinematics algorithm \texttt{singSegInvKin()} \cite{marchese2014design}.
The inputs to this block are the start and endpoint measurements in $\mathbb{R}^2$: $\mathbf{E}_n \; \forall n  = 1..N$, where $N$ is the number of segments composing the arm.
The outputs from this block are the representations of the measured manipulator configuration: \rkk{measured curvature} $\boldsymbol{\kappa}_{\textrm{meas}}$ and \rkk{segment length} $\mathbf{L}_{\textrm{meas}}$.
%\rkk{Each segment of the manipulator is described by its measured curvature $\kappa_{\textrm{meas}}$ and its segment length $L_{\textrm{meas}}$}.
A cascaded closed-loop curvature controller \texttt{curvatureController()} takes the target curvatures $\boldsymbol{\kappa}_{\textrm{target}}$ as its input and resolves the error between $\boldsymbol{\kappa}_{\textrm{target}}$ and $\boldsymbol{\kappa}_{\textrm{meas}}$ by continuously adjusting the fluidic drive cylinder array.

\rkk{We build on the path planning algorithm presented in \cite{marchese2014whole}. This prior work plans the motion of a soft arm without a gripper through a maze at a centerline while taking the arms bulging shape into account. The approach does not work for approaching and grasping objects, since a tip trajectory for successfully moving towards the object is not known, but needs to be generated by posing and solving a new optimization problem. This tip trajectory has to avoid that the manipulator trunk pushes the object prematurely away when approaching it. Therefore, a new planner had to be developed and is presented in the following section.}

\subsection{Autonomous Grasp-and-Place System}
\label{subsec:grasp-place-planner}
The robotic manipulation system is capable of autonomously performing grasp-and-place operations. 
A state flow diagram describing its sensing, planning and executions states is given in Fig. \ref{fig:grasp-and-place-planner}. 
A motion tracker constantly captures the position of the object.
%and passes it along to a routine, which checks for the object to settle.
%\rkk{We consider the object to have settled if it has not moved by more than a small $\epsilon$ for 2\unit{s}}.
%After it has settled,
The Grasp Object Planner receives the coordinates and radius of the object and together with the current curvature values of the arm and gripper, it solves a series of constrained nonlinear optimization problems to generate end-effector poses approaching the object.
Those end-effector poses are waypoints for an optimized path the robot arm should take to get to the final position without the risk of moving the object before the gripper grasps it. 
The intermediate waypoints ensure that the arm moves to the object while its null space maintains a convex shape, \rkk{always} bending away from the object. 
\rkk{This is a conservative approach for
%requiring minimal computation to solve the planning problem of 
not prematurely colliding with the object.} 
Furthermore, \rkk{this approach} allows the arm to move in smaller steps, decreasing the risk of large overshoots due to slip-stick friction between the roller supports and the ground.
This planner is described in more detail in Section~\ref{subsec:grasp_planner}.

\begin{figure}[!htb]
\centering
   \includegraphics[width=0.95\columnwidth]{Figures/processing_control/grasp_place_planner.eps}
   \caption{State flow diagram of the grasp-and-place planner developed for the autonomous grasp-and-place operation of the manipulator. This diagram describes essentially the flow of information from the motion tracking system to the discrete hardware.}
   \label{fig:grasp-and-place-planner}
\end{figure}

The Grasp Object Planner passes the approach configurations $\boldsymbol{\kappa}_i^*$ of the arm to the \texttt{curvatureController()} for execution in real-time.
The controller receives measured curvatures $\boldsymbol{\kappa}_{meas}$ and lengths $\boldsymbol{L}_{meas}$ at an update rate of $100\unit{Hz}$ from the recursively called \texttt{singSegInvKin()} and uses them to successively control the arm to every intermediate configuration $\boldsymbol{\kappa}_i^*$.
During the arm initialization, the new curvature controller performs a pre-pressurization of both lateral channels. 
This is only done for the two segments closest to the root of the arm in order to stiffen them and shorten their response time constant.  
To allow for smoother transitions between each configuration $\boldsymbol{\kappa}_i^*$, we also added a trajectory generation procedure \texttt{trajGen()} to the new curvature controller. It generates in real-time velocity profiles with acceleration and velocity constraints for each individual degree-of-freedom.
These profiles allow real-time interpolation between the approach configurations of the arm while avoiding overshooting at the next target configuration.
When the arm has arrived at the desired pose next to the object, the curvature controller initiates \texttt{graspObject()}.
After encapsulating the object, \texttt{moveToBin()} requests \texttt{trajGen()} for another trajectory from the current pose to a pre-defined bin location.
\rkk{When the manipulator gets close to the bin location}, the procedure \texttt{releaseObject()} causes the gripper to open and release the object.

\subsection{Grasp Object Planner}
\label{subsec:grasp_planner}
A fundamental component of the grasp-and-place system is its ability to plan a feasible approach motion to the object.
That is, given the location $\left(x_o, \, y_o\right)$ and radius $r_o$ of a cylinder enclosing the object as well as the manipulator's current configuration $\boldsymbol{\kappa}_{\textrm{meas}}$ \rkk{and segment lengths $\mathbf{L}_{\textrm{meas}}$}, we determine a series of locally optimal manipulator configurations called approach configurations $\boldsymbol{\kappa}_i^* \; \forall i \, = 1.. numMoves$ that will, if sequentially achieved, bring the manipulator gradually closer to the object while any part of the arm is not touching the object.

The process for determining the approach configurations is detailed in the \texttt{planGrasp()} procedure within the Grasp Object Planner, see Algorithm \ref{alg:planGrasp}.
The planner is visualized in Figure \ref{fig:planGrasp}.
In short, we define a series of approach radii $r_{a_i} \;  \forall i\, = 1.. numMoves$ that define concentric circles shrinking from the manipulator's starting tip pose towards the center of the object.
Given actuator limits, we then search for a series of feasible manipulator configurations $\boldsymbol{\kappa}_i^*$ that will place the robot's end-effector on the approach circles, parameterized by $r_a$ and $\phi$, while minimizing manipulator deformation.
\rkk{Minimized manipulator deformation is chosen as the optimization criterion, because it is proportional to the energy consumed by the fluidic drive cylinders and it also minimizes the strain to the soft actuators.}

\begin{figure}[htpb]
\centering
   \includegraphics[width=0.85\columnwidth, trim = 0mm 0mm 5mm 5mm, clip]{Figures/processing_control/grasp_object_planner.eps}
   \caption{Visualization of the Grasp Object Planner. Concentric approach circles are centered about the object. The locally optimal approach configurations of the arm are shown as curved dashed lines with a straight gripper at the end. The initially measured manipulator configuration is shown as a long and dashed straight line on the right.}
   \label{fig:planGrasp}
\end{figure}

\begin{algorithm}[!htbp]
\begin{small}
  \SetAlgoLined\DontPrintSemicolon
  \KwIn{
  $\boldsymbol{\kappa}_{\textrm{meas}}$, $\mathbf{L}_{\textrm{meas}}$ $\gets$ measured arm configuration \;
  $\boldsymbol{\kappa}_{\textrm{off}}$ $\gets$ \rkk{measured manipulator} configuration at start \;
  $g_{\textrm{off}}$ $\gets$ gripper offset normal to end-effector\;
  $x_o$, $y_o$, $r_o$ $\gets$ object center coordinates and radius \;
  $N$ $\gets$ number of manipulator segments}
  \SetKwFunction{planGrasp}{planGrasp}
  \SetKwFunction{findOptimalConfig}{findOptimalConfig}
  \SetKwFunction{forwKin}{forwKin}
  \SetKwProg{myproc}{Procedure}{}{}
  \myproc{\planGrasp{}}{
  $\mathbf{w}_t \gets \forwKin \left( \boldsymbol{\kappa}_{\textrm{meas}}, \, \mathbf{L}_{\textrm{meas}}, \, N, \, L_N \right)$. \;
  $d_1 \gets \| \left[ x_o, \, y_o \right]^{\textrm{T}} \, - \, \mathbf{w}_t \|$. \;
  $d_2 \gets d_1 - r_o - g_{\textrm{off}}$. \;
  $numMoves \gets \lfloor \frac{d_2}{\Delta d} \rfloor$. \;
  $i$ = 0. \;
  \Repeat{$i$ = $numMoves$}{
    $i$ = $i \, + \, 1$ \;
    $r_{a_i} \gets d_1 \, - \, i\,\frac{d_2}{numMoves}$. \;
    $\boldsymbol{\kappa}_i^* \gets$ \findOptimalConfig{$r_{a_i}$}\;
  }
  \KwRet{$\boldsymbol{\kappa}_i^* \quad \forall i \, = 1.. numMoves$} \;}{}
  \setcounter{AlgoLine}{0}
  \SetKwProg{myproc}{Procedure}{}{}
  \myproc{\findOptimalConfig{$r_{a_i}$}}{
  $\begin{aligned}
    & \boldsymbol{\kappa}^* \, \gets \, \underset{\phi, \, \boldsymbol{\kappa}}{\text{min}}
    & & \mathbf{R} \, \left( \boldsymbol{\kappa} \, - \, \boldsymbol{\kappa}_{\textrm{off}} \right)^2. \\
    & \text{subject to}
    & &  \mathbf{w}_t \gets \begin{bmatrix} x_o \, + \, r_{a_i} \, \cos{\phi} \\ y_o \, + \, r_{a_i} \, \sin{\phi} \\ \phi \, + \, \frac{\pi}{2} \end{bmatrix}. \\
    & & & \mathbf{f} \gets \forwKin\left( \boldsymbol{\kappa}, \, \mathbf{L}, \, N, \, L_N \right). \\
    & & & \mathbf{w}_t - \mathbf{w}_{\textrm{off}}\left(r_o, \, \phi \right) - \mathbf{f} \, = \, \mathbf{0}. \\
    & & & \kappa_{n}^{\text{min}} \leq \kappa_n \leq \kappa_{n}^{\text{max}} \quad \forall \, n = 1 \, .. \,N. \\
    \end{aligned}$ \;
  \KwRet{$\boldsymbol{\kappa}^*$} \; }
  \setcounter{AlgoLine}{0}
  \SetKwProg{myproc}{Procedure}{}{}
  \myproc{\forwKin{$\boldsymbol{\kappa}$, $\mathbf{L}$, $i$, $s$ }}{
  \KwIn{$\boldsymbol{\kappa}$, $\mathbf{L}$, $i$ the segment of interest index, $s$ the arc length along the indexed segment}
  \eIf{$i = 0$}{
        $\theta_{i}(0) \gets \theta_0(0)$. \;
        $x_{i}(0) \gets 0$.\;
        $y_{i}(0) \gets 0$.\; }
    {
        $\left[ x_{i}(0), \, y_{i}(0), \, \theta_{i}(0) \right] \gets \forwKin \left(\boldsymbol{\kappa}, \, \mathbf{L}, \, i-1, \, L_{i-1} \right)$.\;
    }
    $\theta \gets \theta_{i}(0) + k_i s$. \;
    $x \gets x_{i}(0) + \frac{\sin{\theta}}{k_i} - \frac{\sin{\theta_{i}(0)}}{k_i}$. \;
    $y \gets y_{i}(0) - \frac{\cos{\theta}}{k_i} + \frac{\cos{\theta_{i}(0)}}{k_i}$. \;
  \KwRet{$\left[ x, \, y, \, \theta \right]^{\mathrm{T}}$ or $\left[ x, \, y \right]^{\mathrm{T}}$}}
  \caption{Grasp Object Planner}
  \label{alg:planGrasp}
\end{small}

\end{algorithm}

The procedure \rkk{\texttt{planGrasp()} in Algorithm \ref{alg:planGrasp}} first determines the manipulator's current tip pose $\mathbf{w}_t$ and the Euclidean distance $d_1$ between the tip and the object's center. 
\rkk{The arc length input to the arm's forward kinematics is the $N$-th element of the segment lengths $\mathbf{L}_{\textrm{meas}}$.}
\rkk{The end effector offset $\mathbf{w}_{\textrm{off}}$ describes the distance from the root of the gripper to an offset point close to the lower end of the gripper's palm. It is visualized in the top left corner of Fig.~\ref{fig:grasp-and-place-planner}.}
\rkk{The length} $\mathbf{g}_{\textrm{off}}$ represents the component of the end effector offset $\mathbf{w}_{\textrm{off}}$, which is normal to the end effector orientation. 
The minimal tip transit distance $d_2$ is calculated by \rkk{considering the object's radius $r_o$ and the gripper normal offset $\mathbf{g}_{\textrm{off}}$.} 
Also, the number of approach configurations $numMoves$ is determined as $\lfloor \frac{d_2}{\Delta d} \rfloor$, where $\Delta d$ is an allowable incremental distance. 
Using these parameters, approach radii $r_a$ \rkk{shown by the green circles} are iteratively calculated and their corresponding locally optimal configurations are found \rkk{by using the optimization equation and constraints described in procedure} \texttt{findOptimalConfig($r_{a_i}$)} \rkk{of Algorithm \ref{alg:planGrasp}}.

\rkk{The procedure \texttt{findOptimalConfig($r_a$)}} is posed as a nonlinear optimization \rkk{problem}. 
Here, the objective function represents the summation of independently weighted manipulator curvatures \rkk{$\boldsymbol{\kappa} - \boldsymbol{\kappa}_{\textrm{off}}$.}
\rkk{The weights are set by the matrix $\mathbf{R}$.}
\rkk{The variables to optimize for are $\phi$ and $\kappa$.}
The optimization constraints \rkk{cause} the manipulator's tip to lie on and to be tangent to the approach circle. 
\rkk{The constraints also ensure that the manipulator segment curvatures do not exceed the single soft actuator limits.}
\rkk{Furthermore,} this \rkk{procedure} leverages the arm's forward kinematics \texttt{forwKin()} defined in \cite{marchese2014design} and reproduced \rkk{in the last section} of Algorithm~\ref{alg:planGrasp} for convenience. 
\rkk{The optimization only becomes over-constrained if it has to find an arm pose outside of the arm's reachable workspace. That occurs if the object was user-placed outside the workspace. Before performing the optimization, a feasability check is performed using the arm's forward kinematics.}

\rkk{The nonlinear optimization problem was implemented using Matlab’s Optimization Toolbox with the function calls \emph{fmincon}, which finds the minimum of a constrained nonlinear multivariable function. Sequential Quadratic Programming was used as the solver with a relative upper bound of $2 \times 10^{-3}$ on the magnitude of the constraint functions. The lower bound on the size of a step was given by $1 \times 10^{-6}$. The solver takes about 1\unit{s} on a regular PC to solve for all waypoints from start to finish.}