### 2018-04-08 SoRo reviewer comments/2darmgripper source file resubmission/abstract.tex
submission files

This paper presents the development of an autonomous motion planning algorithm for a soft planar grasping manipulator capable of grasp-and-place operations by encapsulation
with uncertainty in the position and shape of the object.
The end effector of the soft manipulator is fabricated in one piece without weakening seams using lost-wax casting instead of the commonly used multi-layer lamination process.
The soft manipulation system can grasp randomly positioned objects within its reachable envelope and move them to a desired location without human intervention.
The autonomous planning system leverages the compliance and continuum bending of the soft grasping manipulator to achieve repeatable grasps in the presence of uncertainty.
A suite of experiments is presented that demonstrates the system's capabilities.